Network Configuration Synthesis with Abstract Topologies

### 2016-11-15 paper/abstraction/paper.tex
a few edits

To configure large networks, instead of considering individual
devices, operators classify devices into {\em roles}. A role refers to
specific functionality and is served by one or more devices. For instance, in a
data center, roles may be ``top-of-rack,'' ``aggregation,'' and ``spine''
routers; and in a backbone network, they may be ``core'' and ``border''
routers. While the network may have hundreds or thousands of devices---a scale
that is impossible for humans to handle---there tend to be a handful of roles.
%
Operators author a configuration {\em template} for each role. Templates are
macros that, given a network topology, can be instantiated with different concrete values to generate device configurations.

@@

Based on these inputs, \sysname generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the resulting
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done in the abstract domain by using a set of sound inference rules to infer the lower bound on the number of edge-disjoint paths in any concrete topology.
%Our analysis uses a set of sound inference rules and  using an optimizing SMT solver.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of the \sysname tool that generates
provably-correct templates and is much faster
than tools based on concrete topologies.
\end{itemize}

@@

As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy ($(8)$) when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

More specifically, a network topology is a graph $G$ = ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$. A role-based abstraction is a graph homomorphism from $G$ to an abstract graph $G^A$ = ($V^A$,$E^A$). A graph homorphism $f : G \rightarrow G^A$ maps each node in the concrete graph to a node in the abstract graph such that, whenever $(u,v) \in E$, then $(f(u),f(v)) \in E^A$. The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs.

@@

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local constraint applies to any traffic destined for a local prefix described by the variable $\CD{\$LP}$ and adds the constraint that the traffic must follow a path that matches an internal location (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each data center level.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

\para{Fault-tolerance policy} This policy specifies how many link failures the network should be able to withstand before traffic experiences connectivity loss. Operators can specify different tolerance levels for different pairs of abstract nodes. For instance, they may say that ToR to spine connectivity should be robust to 2 failures, \IE, no ToR-spine pair should lose connectivity as long as the number of simultaneous link failures is 2 or fewer; and ToR-to-ToR connectivity should be be robust to 1 failure.

@@

To learn about disjoint paths, we develop a collection of sound inference rules for invariants of the form above. Figure~\ref{fig:inference-rules} shows each rule. They are read from bottom to top. The label on the bottom left is a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to invariants learned after applying the inference rule. The adjacent box shows the conditions that must be \emph{valid} for the rule to apply.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is useful particularly useful for data center topologies where routers in one tier of the data center often have a particular striping pattern with another tier.

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. This ensures that the condition holds for all concrete instantiations of the topology.  Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Invariants learned with $j=0$ or $k=0$ are discarded.

>>>>

To configure large networks, instead of considering individual
devices, operators classify devices into {\em roles}. A role refers to
specific functionality and is served by one or more devices. For instance, in a
data center, roles may be ``top-of-rack,'' ``aggregation,'' and ``spine''
routers; and in a backbone network, they may be ``core'' and ``border''
routers. While the network may have hundreds or thousands of devices---a scale
that is impossible for humans to handle---there tend to be only a handful of roles.
%
Operators author a configuration {\em template} for each role. Templates are
macros that, given a network topology, can be instantiated with different concrete values to generate device configurations.

@@

Based on these inputs, \sysname generates one template per role. It
guarantees that, when the templates are instantiated to configurations for {\em
any} concrete topology that complies with the abstract topology, the resulting
network will correctly implement the routing policy and meet the fault-tolerance
requirements. We show how fault-tolerance analysis can be done in the abstract domain by using a set of sound inference rules to infer a lower bound on the number of edge-disjoint paths in any concrete topology.
%Our analysis uses a set of sound inference rules and  using an optimizing SMT solver.

@@

\vspace{1em}
\noindent
\textbf{Contributions}: To summarize, our contributions are:
%
\begin{itemize}[noitemsep,nolistsep,label={\large\textbullet}]
\item New topology abstractions for network programming
based on graph homomorphisms and connectivity rules.
\item New algorithms for analyzing routing policy and fault tolerance over
abstract topologies.
%one which guarantees correct compilation and one which determines fault tolerance
%properties of network policy.
%Both algorithms provide guarantees for any
%concrete network satisfying the abstract connectivity invariants.
%\item Design of a template generation strategy that supports network
%evolution.
\item Implementation and evaluation of the \sysname compiler that generates
provably-correct templates and is much faster
than compilers based on concrete topologies.
\end{itemize}

@@

As described earlier (\S\ref{sec:introduction}), templates are hard to author and hard to validate. Worse, templates that work for one topology may not work for seemingly-inconsequential variations (which may arise after the network evolves).
Consider the network in Figure~\ref{fig:example2}, which is similar to Figure~\ref{fig:example}: it has the same five roles, connected in a similar hierarchy. One might think that the same templates (with different database entries) can be used for both cases. However, if the templates are configured to disallow ``valley" paths (per policy (5) above), they will work for Figure~\ref{fig:example} but silently violate the fault tolerance policy $(8)$ when used for Figure~\ref{fig:example2}.  Specifically, in Figure~\ref{fig:example2}, an aggregation-based blackhole (\S\ref{sec:background}) will occur when the link S1--A1 fails; after this failure, S1 has no valley-free path to P[1--2] even though it will continue to get traffic for these prefixes as it announces the covering prefix PG (because it gets routes for P[3--4]). No such blackhole will occur in Figure~\ref{fig:example} after a single link failure because spine routers have two links to each pod.

@@

More specifically, a network topology is a graph $G$ = ($V, E$), which consists of a set of vertices $V$ and a set of directed edges $E \colon V \times V$. A role-based abstraction is a graph homomorphism from $G$ to an abstract graph $G^A$ = ($V^A$,$E^A$). A graph homomorphism $f : G \rightarrow G^A$ maps each node in the concrete graph to a node in the abstract graph such that, whenever $(u,v) \in E$, then $(f(u),f(v)) \in E^A$. The role-based abstraction therefore overapproximates the connectivity of the underlying concrete graphs.

@@

Next we can capture constraint (4) that traffic for local prefixes must stay within in the data center:
%
\begin{code}
\Define Local =
    \$LP \Path \Always(\In)
\end{code}
\noindent%
%
The Local policy adds the $\CD{\Always(\In)}$ constraint for each local prefix described by the template variable $\CD{\$LP}$. This constraint ensures that traffic follows a path that matches an internal location (\IE, in the data center) at each hop of the path.
%
The constraint to prevent ``valleys" (5) is written as:
%
\begin{code}
\Define NoValley =
    \True \Path \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent%
%
This policy applies to all traffic and prevents valley paths by adding the $\Novalley$ constraint with arguments corresponding to each level in the data center.
%
Constraint (7) to prevent transit traffic between neighbors is expressed as:
%
\begin{code}
\Define Peer = \{N1,N2\}
\Define NoTransit =
    \True \Path !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent%
%
We define a $\CD{Peer}$ as $\CD{N1}$ or $\CD{N2}$ and disallow paths where traffic both enters and exits the data center through a peer.

@@

\para{Fault-tolerance policy} This policy specifies how many link failures the network should be able to withstand before traffic experiences connectivity loss. Operators can specify different tolerance levels for different pairs of abstract nodes. For instance, they may say that ToR to spine connectivity should be robust to 2 failures, \IE, no ToR-spine pair should lose connectivity as long as the number of simultaneous link failures is 2 or fewer; and ToR-to-ToR connectivity should be robust to 1 failure.

@@

To learn about disjoint paths, we develop a collection of sound inference rules for invariants of the form above. Figure~\ref{fig:inference-rules} shows each rule. They are read from bottom to top. The label on the bottom left is a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to invariants learned after applying the inference rule. The adjacent box shows the conditions that must be \emph{valid}, given the abstraction constraints, for the rule to apply. Validity ensures that the condition holds for all concrete instantiations of the topology.

@@

The second part of the rule uses a similar idea to reason about the overlap between roles $m$ and $o$ with respect to role $n$. This rule is particularly useful for data center topologies where routers in one tier of the data center often have a particular striping pattern with another tier.

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any invariant learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Invariants learned with $j=0$ or $k=0$ are discarded.
